My title

### 2024-11-03 paper/paper.qmd
model update with vizualisation

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

#Load and clean data
data <- read.csv("../data/01-raw_data/raw_data.csv") |>
  clean_names()
```

# Select imporntat variables (change as you please)

preddata <- data %>%
  filter(numeric_grade >1) %>%
  select(pollster, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date, population, pollscore, numeric_grade) %>%
  filter(!is.na(pollscore)) %>%
  filter(!is.na(state) & state != "")

print(preddata)

#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?
harrisdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    end_date = mdy(end_date)    #is date mutated correctly here
  ) %>%
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(
    num_harris = round((pct / 100) * sample_size, 0)
  )

harrisdata

# Plots poll estimates and overall smoothing
base_plot +
  geom_point() +
  geom_smooth()

# This gets messy - need to add a filter - see line 21
base_plot +
  geom_point(aes(color = state)) +
  geom_smooth() +
  theme(legend.position = "bottom")

## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE AND STATE

# Define Model 1
model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)

bayesian_model <- stan_glmer(
  formula = model,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

# Plot checks for Model
pp_check(bayesian_model)
summary(bayesian_model)
plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
print(modelsummary(bayesian_model))
summary(bayesian_model)
```

# Predict and plot

spline_model_pred <- stan_glm(
  pct ~ ns(end_date_num, df = 5) + state + pollster,
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

# Summarize the model
summary(spline_model_pred)

# Posterior predictive checks
pp_check(spline_model_pred)
```

# Convert `pollster` to a factor with defined levels if it's not already
harrisdata$pollster <- factor(harrisdata$pollster, levels = unique(harrisdata$pollster))

# Define the new data grid for prediction, ensuring each factor has the correct levels
new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100
  ),
   pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
  state = factor("Georgia", levels = levels(harrisdata$state)))

posterior_preds <- posterior_predict(spline_model_pred, newdata = new_data)

ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit") +
   facet_wrap(~ state, scales = "free_y") +
  theme_minimal()
```

@@

As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election. Our original Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll. Each poll has two entries one giving the Data for the polling a opinions of Donald Trump and the second for Kamala Harris. We have acquired our polling data from @bibnotmadeyet.

@@

**display_name:** This column shows the name of the Pollster that conducted the poll, as we want to keep consistency we are only looking at polls conducted by the New York Times.

**question_id:** This variable refers to the question that was asked in the specific poll.

**sample_size:** This variable is the quantity of people that participated in the specific poll.

**state:** This variable tells us in which States the poll was conducted.

**party:** This variable tells us which party the poll participants intend to vot for.

**answer:** This variable tells us the candidate the participant intend to vote for.

**candidate_id:** This variable is the numerical value assigned to each candidate.

**candidate_name**: This is the full name of the selected candidate.

**pct:** This tells us the percentage of participants that intend to vote for the selected candidate.

**start_date:** The date the pollster began conducting the poll.

**end_date:** The date the pollster finished conducting the poll.

**num_harris:** The number of participants that intend to vote for Kamala Harris.

**end_date_num:** IDK

@@

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
```
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

# Discussion

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/Mezhi18/US_Election2024.git](https://github.com/Mezhi18/US_Election2024.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    bibliography: references.bib
    documentclass: article
    geometry: margin = 1in
number-sections: true
toc: true
bibliography: references.bib
---

@@

@@

As our paper is about the 2024 United States federal election and more specifically we are looking at the polls and the polling data comparing the two candidates, former President Donald Trump and Vice President Kamala Harris, for the upcoming election. Our original Data set had over 16,000 unique entries from different pollsters, the business or Organization that conducts the poll. Each poll has two entries, one giving the Data for the polling opinions of Donald Trump and the second for Kamala Harris. We have acquired our polling data from @bibnotmadeyet.

@@

- `pollster`: Shows name of the Pollster that conducted the poll.
- `sample_size`: The number of people that participated in the specific poll.
- `state`: This variable tells us in which States the poll was conducted.
- `candidate_name`: This is the full name of the selected candidate.
- `pct`: This tells us the percentage of participants that intend to vote for the selected candidate.
- `start_date`: The date the pollster began conducting the poll.
- `end_date`: The date the pollster finished conducting the poll.

The variables that we have create are:

- `num_harris`: The number of participants that intend to vote for Kamala Harris.
- `end_date_num`: The number of days since the first poll since Harris announced her candidacy.

Each pollster has a numeric grade from 1.0 to 3.0, which indicates the quality/ reliability of the respective pollster. Additionally, each pollster is also given a transparency score from 1.0 to 10.0 reflecting how 'transparent' the pollster is, or how much information is disclosed about its polls and methodology. It is important for pollsters to maintain high numeric grades and transparency scores because these metrics directly reflect the quality and reliability of their data. To ensure the highest level of accuracy in our predictions, we only include polls with a numeric grade of 1.5 or above and a transparency score of 6.0 or above.

¿ does this go into data cleaning ?

@@

## Methodology

This paper uses 'polls-of-polls' method to analyze and predict our outcome, which combines polls from multiple sources (pollsters). By combining results from various sources, this approach incorporates diverse perspectives, which helps to minimize individual biases and provides a more balanced view. Unlike relying on a single poll, which can be influenced by its own biases, the 'polls-of-polls' method enhances reliability and increases the overall validity of the results.

... explain how pollsters survey people, what their different methodologies are, and why these things are important for an accurate prediction...

## Measurement/ Data Visualization

#Load and clean data
data <- read.csv("../data/01-raw_data/raw_data.csv") |>
  clean_names()

preddata <- data %>%
  filter(numeric_grade >1.5 & transparency_score >6) %>%
  select(pollster, sample_size, state, candidate_name, pct, start_date, end_date, pollscore, numeric_grade) %>%
  filter(!is.na(pollscore)) %>%
  filter(!is.na(state) & state != "")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-supportstate
#| fig-cap: "Distribution of Support Percentage by State"

harrisdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(end_date = mdy(end_date)) %>%
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(num_harris = round((pct / 100) * sample_size, 0))

ggplot(harrisdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
  geom_boxplot() +
  coord_flip() +  # Flip for better readability
  labs(x = "State", y = "Support Percentage") +
  theme_minimal() +
  theme(legend.position = "none")
```
We focus on the swing states of the U.S, according to  [cite NYT PAPER]

These are the statistics of each swing state by percetnage by state.

base_plot +
  geom_point(aes(color = state)) +
  geom_smooth() +
  theme(legend.position = "bottom")

ggplot(harrisdata, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_smooth(se = TRUE) +
  facet_wrap(~ state, scales = "free_y") +
  labs(y = "Harris Percent", x = "Date", title = "Polling Trends for Kamala Harris by State") +
  theme_minimal()

@@

@@

@@

This is Model 1, follows Bayesian Model

We define our model as:

*define each vairable here*
```{=tex}
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{State}_i + \beta_2 \times \mbox{Days Since First Poll}_i + \beta_3 \times \mbox{Pollster}_i \\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Exponential}(1)
\end{align}
```

# Define Model
model <- cbind(num_harris, sample_size - num_harris) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)

priors <- normal(0, 2.5, autoscale = TRUE)

bayesian_model <- stan_glmer(
  formula = model,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  weights = harrisdata$numeric_grade,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

# Plot checks for Model
pp_check(bayesian_model)
plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
modelsummary::modelsummary(
  list(
    "First model" = bayesian_model
  ),
  statistic = "mad",
  fmt = 2
)
```

This is Model 2

spline_model <- stan_glm(
  pct ~ ns(end_date_num, df = 5) + pollster + state, 
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

# Summarize the model
summary(spline_model)

# Posterior predictive checks
pp_check(spline_model)

modelsummary::modelsummary(
  list(
    "First model" = spline_model
  ),
  statistic = "mad",
  fmt = 2
)
```

harrisdata$pollster <- factor(harrisdata$pollster)
harrisdata$state <- factor(harrisdata$state)

new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100
  ),
  pollster = factor("YouGov", levels = levels(harrisdata$pollster)),
  state = factor("Georgia", levels = levels(harrisdata$state)))

posterior_preds <- posterior_predict(spline_model, newdata = new_data)

# Plot data and Bayesian model predictions with confidence ribbon
ggplot(harrisdata, aes(x = end_date_num, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE
  ) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time"
  ) +
  facet_wrap(~ state, scales = "free_y") +
  theme_minimal()
```

# TRUMP #

```{r}
trumpdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "Wisconsin")) %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(end_date = mdy(end_date)) %>%
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(num_trump = round((pct / 100) * sample_size, 0))

ggplot(trumpdata, aes(x = reorder(state, pct, median), y = pct, fill = state)) +
  geom_boxplot() +
  coord_flip() +  # Flip for better readability
  labs(x = "State", y = "Support Percentage") +
  theme_minimal() +
  theme(legend.position = "none")

trumpdata <- trumpdata %>%
  mutate(
    end_date_num = as.numeric(end_date - min(end_date)))

trumpdata <- trumpdata |>
  mutate(state = factor(state))

# Define Model
model2 <- cbind(num_trump, sample_size - num_trump) ~ (1 | state) + (1 | pollster) + (1 | end_date_num)

priors <- normal(0, 2.5, autoscale = TRUE)

bayesian_model2 <- stan_glmer(
  formula = model2,
  data = trumpdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  weights = trumpdata$numeric_grade,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95)
```

# Plot checks for Model
pp_check(bayesian_model2)
plot(bayesian_model2, pars = "(Intercept)", prob = 0.95)
modelsummary::modelsummary(
  list(
    "First model" = bayesian_model2
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

@@

What do we expect...
We expect a predictions in the high 40% based on our current data.

@@

```{r}
harrisdata <- harrisdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_harris))
trumpdata <- trumpdata %>% filter(!is.na(numeric_grade)) %>% filter(!is.na(num_trump))
trumpdata
```

```{r}
library(dplyr)
library(ggplot2)
library(rstanarm)

harris_data <- harrisdata %>%
  mutate(
    candidate_name = "Kamala Harris",
    support_ratio = num_harris  # Assuming harris_support_ratio is already calculated
  ) %>%
  select(end_date_num, candidate_name, support_ratio)

# Prepare Trump data with a unified structure
trump_data <- trumpdata %>%
  mutate(
    candidate_name = "Donald Trump",
    support_ratio = num_trump  # Assuming trump_support_ratio is already calculated
  ) %>%
  select(end_date_num, candidate_name, support_ratio)

# Combine both into a single data frame
combined_data <- bind_rows(harris_data, trump_data)

library(ggplot2)
library(ggplot2)

# Assuming `combined_data` contains columns `end_date_num`, `support_ratio`, and `candidate_name`

ggplot(combined_data, aes(x = end_date_num, y = support_ratio, color = candidate_name, fill = candidate_name)) +
  # Shaded area around each line for overlap visualization
  geom_ribbon(data = combined_data %>% filter(candidate_name == "Kamala Harris"),
              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
              alpha = 0.3, fill = "red") +
  geom_ribbon(data = combined_data %>% filter(candidate_name == "Donald Trump"),
              aes(ymin = support_ratio - 0.05, ymax = support_ratio + 0.05),
              alpha = 0.3, fill = "blue") +

  # Line plot for each candidate
  geom_line(aes(group = candidate_name), size = 1) +

  scale_color_manual(values = c("Kamala Harris" = "red", "Donald Trump" = "blue")) +  # Set line colors
  labs(
    x = "Days Since Earliest Poll", 
    y = "Support Ratio",
    color = "Candidate",
    title = "Support Ratio for Harris and Trump Over Time with Overlap Highlighted"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    legend.title = element_text(face = "bold")
  )

```
# Discussion

@@

```{r}
#| eval: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

@@

```{r}
#| echo: false
#| eval: false
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

@@

# Appendix 1

# Appendix 2

=======
>>>>>>> eda40e1517683f0ad12e0bcd901797a2bb2abd7d
\newpage
